\documentclass{acmsmall}

\usepackage{scalerel}
\usepackage{import}
\usepackage{times}
\usepackage{multirow}
\usepackage{microtype}
\usepackage{array}
\usepackage{galois}
\usepackage{graphicx,wrapfig}
\usepackage[noadjust]{cite}
\usepackage{tikz}
\usetikzlibrary{plotmarks}
\usepackage{pgfplotstable}
\usepackage{filecontents}
\usepackage{pgfplots}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{psfrag}
\usepackage{epsfig}
\usepackage{wasysym} 
%\usepackage{subfigure}
\usepackage{paralist}
%\usepackage{dingbat}
\usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
\usepackage{hyperref}
\usepackage[subnum]{cases}
\makeatother

\newcommand{\extendedonly}[1]{}
\newcommand{\paperonly}[1]{#1}
\newtheorem{exmp}{Example}
%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}[theorem]{Lemma}
%\numberwithin{mytheorem}{subsection} 

\newcommand{\tool}[1]{\textsc{#1}\xspace}
\newcommand{\cbmcv}{\tool{cbmc 5.0}}
\newcommand{\mydef}[1]{\begin{definition}#1\end{definition}}
\newcommand{\rmcmt}[1]{{\color{magenta}{#1}}}%#1
\newcommand{\pscmt}[1]{{\color{blue}{#1}}}%#1
\newcommand{\dkcmt}[1]{{\color{red!70!black}{#1}}}%#1
\newcommand{\tmcmt}[1]{{\color{orange}{#1}}}%#1
\newcommand{\lhcmt}[1]{{\color{purple}{#1}}}%#1

\newif\ifextended
\extendedfalse        % For FM'16 publishers version

\input{symbols}

\begin{document}

% ---------------------------------------------------------------------
% Title, authors, abstract, keywords
% ---------------------------------------------------------------------
\title{Precise Abstract Interpretation Using CDCL}

\author{Rajdeep Mukherjee \and Peter Schrammel \and 
Leopold Haller \and \\ 
Daniel Kroening \and Tom Melham}

%\authorrunning{Mukherjee, Schrammel, Haller, Kroening, Melham}

%\institute{University of Oxford, UK \and University of Sussex, UK \and Google Inc., USA}

\maketitle

%===============================================================================
\begin{abstract}
%===============================================================================
%
  This paper shows that Conflict Driven Clause Learning (CDCL) algorithm
  employed in modern day satisfiability solvers can be used to compute 
  fixed point approximations over lattice of program traces to determine 
  safety.  Satisfiability or model finding can be seen as a property of fixed 
  points of unsafe trace transformers over this lattice.  CDCL algorithm 
  alternates between an overapproximate model search phase and an 
  underapproximate conflict analysis phase.  We show that 
  the model search computes a greatest fixed point over unsafe trace transformer
  using deductions and decisions.  Whereas, conflict analysis computes a 
  least fixed point of safe trace transformer over a downset lattice with 
  a heuristic choice of conflict reasons.  Thus, CDCL architecture can be seen 
  as a natural tool to build a precise program analyzer that uses decision and 
  learning to automatically refine the precision of an analysis. This paper
  present a sound mathematical framework for building a precise abstract 
  interpretation based program analyzer using CDCL architecture as well as a practical tool 
  that implements these ideas for automatic bounded safety verification of C programs. 
  We call our technique \emph{Abstract Conflict Driven Learning for Programs} (ACDLP).
  We~evaluate the performance of our analyser by comparing with CBMC, which
  uses Boolean CDCL, and Astr{\'e}e, a commercial abstract interpretation tool. 
  We~observe two orders of magnitude reduction in the number of decisions,
  propagations, and conflicts as well as a~1.5x speedup in runtime compared to
  CBMC.  Compared to Astr{\'e}e, ACDLP solves twice as many benchmarks and has
  much higher precision.  
  %Our work is the first step towards lifting CDCL algorithm for design and 
  %implementation of a precise static analyzers.
%
\end{abstract}

%===============================================================================
\input{introduction}~\label{intro}
%===============================================================================

%===============================================================================
\input{definition}
%===============================================================================

%===============================================================================
\input{solver}
%===============================================================================

%===============================================================================
\input{program}
%===============================================================================

%===============================================================================
\input{acdlp}
%===============================================================================

%===============================================================================
%\input{completeness}
%===============================================================================

%===============================================================================
\input{proof}
%===============================================================================

%===============================================================================
%\input{ssa}
%===============================================================================

%===============================================================================
%\input{acdlp}
%\input{deductions}
%===============================================================================

%===============================================================================
%\input{sat-example}
%===============================================================================

%===============================================================================
\input{experiments}
%===============================================================================

%===============================================================================
\input{related_work}
%===============================================================================

%===============================================================================
\section{Conclusion}
%===============================================================================
%
The lattice theoretic account of CDCL algorithm provides a natural extension 
to other domains.  In this paper, we present a full abstract interpretation
account of bounded model checking using CDCL architecture.  To this end, we
first show that finding non-trivial trace abstraction in presence of loops, 
that satisfy properties of CDCL can be achieved through logical encoding of
program semantics using static single assignment form.  We then present a
characterization of CDCL as a tool to compute fixed point approximation of unsafe 
trace transformer over a lattice of program traces.  We showed that decision 
and learning in CDCL automatically allows the analysis to recover from
imprecision.  Our framework provides a mathematical basis for instantiating 
CDCL architecture for program analysis using the framework of abstract interpretation. 
A practical benefit of our work is a new, learning based architecture for implementing 
precise \rmcmt{program or safety?} analysis tools, through combination of abstraction
interpretation and CDCL-based SAT Solvers.  

\bibliographystyle{splncs03}
\bibliography{biblio}

\end{document}
